---
title: "Untitled"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(jagsUI)

rm(list = ls())

source('/home/ben/linkedModels/movementModel/moveModelWithData/moveModFunctions.R')

# for now unitl it's in the package
source('/home/ben/linkedModels/movementModel/moveModelWithData/WB/addSectionProperties.R')

## defining data section ##

moveDir <- "/home/ben/linkedModels/movementModel/moveModelWithData/WB/"
setwd(moveDir)

load(file = '/home/ben/linkedModels/dataForModels/cdForMovementModelWB.RData')

###
#prepare and filter the data
cdWB <- cdWB %>% filter( enc == 1 )
cdWB <- filter(cdWB,area %in% c("inside","trib"), section > 0)

#
cdWB <- addSectionProperties(cdWB) #adds in section widths. will be in the getWBData package once Evan gets back to me and will go in getAndPrepareWBData


#limit data to more than one obs per fish
counts <- cdWB %>% group_by(tag) %>% summarize(n=n()) %>% arrange(desc(n))
cdWB <- left_join(cdWB,counts) %>%
        filter(n > 1)  # remove single observations 

cdWB <- cdWB %>% 
#   select(tag,cohort,detectionDate,sampleIndex,river,section,area,observedLength)
   group_by(tagIndex) %>%
   mutate( 
           lagRiver = lead(as.numeric(riverOrdered)),trans=paste0(as.numeric(riverOrdered),lagRiver)
         )

cdWB$isYOY <- ifelse(cdWB$cohort == cdWB$year,1,0)
cdWB$riverMoved <- ifelse(as.numeric(cdWB$riverOrdered) == cdWB$lagRiver,0,1)

```

##Join in mean biomass for current (biomassFrom) and next section (biomassTo)
Testing whether fish move from high biomass sections and into low density sections
```{r}

bC <- cdWB %>% 
  group_by(river,season,year,section) %>%
  summarize( biomass = mean(observedWeight,na.rm=T),
             count = n() ) %>%
  select(biomass,count,river,+season,year,section)

cdWB <- left_join(cdWB,bC) %>%
        rename(biomassFrom = biomass, countFrom = count) 
cdWB <- left_join(cdWB,bC, by=c("lagSection" = "section","river", "year", "season")) %>%
        rename(biomassTo = biomass, countTo = count)


cdWB$moved01 <- ifelse(cdWB$section == cdWB$lagSection, 0, 1)
# to move (=1) must move more than 1 section away
cdWB$moved01PM <- ifelse(cdWB$section == cdWB$lagSection - 1 |
                          cdWB$section == cdWB$lagSection     |
                          cdWB$section == cdWB$lagSection + 1, 0, 1)

#cdWB$sectionFactor <- factor( cdWB$section )
```

In some initial analyses, fish seem to move to high biomass sections. Maybe these are just good sections that can carry high biomass. Try scaling within sections by average biomass over the study as a crude measure of habitat quality

```{r}

ggplot(filter(cdWB), aes(as.factor(section), observedWeight)) +
  geom_boxplot(outlier.size = 0.2) + 
  scale_y_log10() +
  facet_grid(river+isYOY~season, scales = "free")

# means by isYOY,river,year,season,section
# this will be our 'raw' mean biomass and counts
byYRYSS <- cdWB %>% 
  group_by(isYOY,river,year,season,section) %>%
  summarize( meanBiomassYRYSS = mean(observedWeight,na.rm = T),
             meanBiomassByAreaYRYSS = mean(observedWeight/(widthToUse + 0.01), na.rm = T), 
             varBiomassByAreaYRYSS = var(observedWeight/(widthToUse + 0.01), na.rm = T), 
             nYRYSS = n(),
             meanWidthToUse = mean(widthToUse + 0.001, na.rm = T)) %>%
  select(meanBiomassYRYSS,nYRYSS,meanBiomassByAreaYRYSS,varBiomassByAreaYRYSS,meanWidthToUse,isYOY,river,season,year,section)

#byYRYSS$meanBiomassByAreaYRYSS0 <- ifelse(is.nan(byYRYSS$meanBiomassByAreaYRYSS),NA,byYRYSS$meanBiomassByAreaYRYSS)

hist(byYRYSS$meanBiomassByAreaYRYSS)
hist(byYRYSS$meanWidthToUse)
filter(byYRYSS,meanWidthToUse <0.25)
filter(cdWB,meanWidthToUse < 0.1)#%in% c(0.03230321,0.01))
table(cdWB$meanWidthToUse)
```
##make some  plots of our means
```{r}
ggplot(byYRYSS, aes(meanBiomassByAreaYRYSS, meanBiomassYRYSS, color = factor(year))) +
  geom_point() +
  facet_grid(section~river)

ggplot(byYRYSS, aes(section, meanBiomassYRYSS, color = factor(year))) +
  geom_point() +
  geom_line() +  
  facet_grid(river+isYOY~season, scales = "free")

ggplot(byYRYSS, aes(section, nYRYSS, color = factor(year))) +
  geom_point() +
  geom_line() +  
  facet_grid(river+isYOY~season, scales = "free")
```

```{r}
# standardize within sections across years
# using ideas from the 'many models' section of 'R for Data Science' book
### run separate linear models for each isYOY,river,season
sectionModelB <- function(df) {
#  lm( meanBiomassYRYSS ~ factor(section), data = df )
  lm( meanBiomassByAreaYRYSS ~ factor(section), data = df )
}

sectionModelN <- function(df) {
  lm( nYRYSS ~ factor(section), data = df )
}


byYRYSS2 <- byYRYSS %>% 
  #filter(section %in% 1:15) %>%
  filter( !(isYOY == 1 & season %in% 1:2) ) %>% # data are too sparse
#  filter( !(isYOY == 0 & season %in% 4 & river == 'wb mitchell'), year %in% 2005:2007 ) %>% # data are too sparse before Todd's update
#  filter( section<15) %>%
  filter( !is.nan(meanBiomassByAreaYRYSS)) %>% #otherwise end up with 'extra sections error' from map()
  group_by(isYOY,river,season) %>%
  nest() 

tmp <- byYRYSS2 %>% unnest() %>%
  group_by(isYOY,river,season,section) %>%
  mutate( count = n())

#tally(tmp)

#data.frame(byYRYSS2$data[[7]])
#byYRYSS2$modelB[[2]]

# run each model and add residuals
byYRYSS2 <- byYRYSS2 %>% 
  mutate( modelB = map(data, sectionModelB),
          residBiomass = map2(data, modelB, add_residuals),
          modelN = map(data, sectionModelN),
          residN = map2(data, modelN, add_residuals)
      )

# to long format
rB <- unnest(byYRYSS2,residBiomass) %>% rename(residB = resid)
rN <- unnest(byYRYSS2,residN) %>% rename(residN = resid)

ggplot(rB, aes(section,residB, color = factor(year))) +
 geom_line() +
 facet_grid(river+isYOY~season, scales = "free")

ggplot(rN, aes(section,residN, color = factor(year))) +
 geom_line() +
 facet_grid(river+isYOY~season, scales = "free")

```

##Merge biomass and N residuals back into cdWB
```{r}
#cdWB <- left_join(cdWB,select(r, isYOY,river,season,section,residBiomass,residN))
# merge in biomass and counts for the current sample 
cdWB <- left_join(cdWB,rB, by = c("river", "section", "year", "season", "isYOY"))
cdWB <- left_join(cdWB,select(rN, -meanBiomassYRYSS,-nYRYSS,-meanBiomassByAreaYRYSS,-varBiomassByAreaYRYSS,-meanWidthToUse), by = c("river", "section", "year", "season", "isYOY"))

# merge in biomass and counts for the next sample
# need to rename residX so join doesn't rename variables by adding a .x or .y
rBNext <- rB %>% rename(residBNext = residB, meanBiomassYRYSSNext  = meanBiomassYRYSS, nYRYSSNext  = nYRYSS )
cdWB <- left_join(cdWB,rBNext, by = c("lagSection" = "section","river", "year", "season","isYOY")) 
        
rNNext <- rN %>% rename( residNNext = residN )
cdWB <- left_join(cdWB,select(rNNext, -meanBiomassYRYSS,-nYRYSS), by = c("lagSection" = "section","river", "year", "season","isYOY"))

```        

```{r}

ggplot(filter(cdWB,residB<100,residB>-100), aes(residB,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-20,21)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(species~river, scales="free")

summary( glm( moved01 ~ residB, family = "binomial", data = cdWB ) )
summary(lm(residB ~ 0+factor(moved01)*species*river, data=cdWB))

ggplot(cdWB, aes(residB,color=factor(moved01))) +
  geom_freqpoly() +
#  xlim(c(-40,40)) +
  facet_grid(species~river)

#t.test(residB ~ 0+factor(moved01), data=cdWB)

ggplot(cdWB, aes(residBNext,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(species~river, scales="free")

summary( glm( moved01 ~ residBNext, family = "binomial", data = cdWB ) )


ggplot(cdWB, aes(residN,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) 

summary( glm( moved01 ~ residN, family = "binomial", data = cdWB ) )

ggplot(cdWB, aes(residNNext,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) 

summary( glm( moved01 ~ residNNext, family = "binomial", data = cdWB ) )


```


