---
title: "Untitled"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(jagsUI)

rm(list = ls())
source('/home/ben/linkedModels/movementModel/moveModelWithData/moveModFunctions.R')

## defining data section ##

moveDir <- "/home/ben/linkedModels/movementModel/moveModelWithData/WB/"
setwd(moveDir)

load(file='/home/ben/linkedModels/dataForModels/cdForMovementModelWB.RData')

###
#prepare and filter the data
cdWB <- cdWB %>% filter( enc == 1 )

#limit data to more than one obs per fish
counts <- cdWB %>% group_by(tag) %>% summarize(n=n()) %>% arrange(desc(n))
cdWB <- left_join(cdWB,counts) %>%
        filter(n > 1)  # remove single observations 

cdWB <- cdWB %>% 
#   select(tag,cohort,detectionDate,sampleIndex,river,section,area,observedLength)
   group_by(tagIndex) %>%
   mutate( 
           lagRiver = lead(as.numeric(riverOrdered)),trans=paste0(as.numeric(riverOrdered),lagRiver)
         )

cdWB$isYOY <- ifelse(cdWB$cohort == cdWB$year,1,0)
cdWB$riverMoved <- ifelse(as.numeric(cdWB$riverOrdered) == cdWB$lagRiver,0,1)

```

##Join in mean biomass for current (biomassFrom) and next section (biomassTo)
Testing whether fish move from high biomass sections and into low density sections
```{r}

bC <- cdWB %>% 
  filter(area %in% c("inside","trib")) %>%
  group_by(river,season,year,section) %>%
  summarize( biomass = mean(observedWeight,na.rm=T),
             count = n() ) %>%
  select(biomass,count,river,+season,year,section)

cdWB <- left_join(cdWB,bC) %>%
        rename(biomassFrom = biomass, countFrom = count) 
cdWB <- left_join(cdWB,bC, by=c("lagSection" = "section","river", "year", "season")) %>%
        rename(biomassTo = biomass, countTo = count)

#cdWB$biomassFrom <- log(cdWB$biomassFromRaw)
#cdWB$biomassTo <- log(cdWB$biomassToRaw)
cdWB$moved01 <- ifelse(cdWB$section == cdWB$lagSection, 0, 1)
#cdWB$moved01PM <- ifelse(cdWB$section %in% c(cdWB$lagSection - 1, cdWB$lagSection, cdWB$lagSection + 1), 0, 1)
cdWB$moved01PM <- ifelse(cdWB$section == cdWB$lagSection - 1 |
                          cdWB$section == cdWB$lagSection     |
                          cdWB$section == cdWB$lagSection + 1, 0, 1)

```

```{r}

d <- select(cdWB, tag,species,river,lagRiver,sampleIndex,year,season,section,lagSection,riverMoved,moved01,moved01PM,distMoved,biomassFrom,biomassTo )

```

##All the data
```{r}

# ggplot(d, aes(biomassFrom,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) 
# 
# ggplot(d, aes(biomassTo,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) 
```

##Plot by spp
```{r, warning=FALSE}

# ggplot(d, aes(biomassFrom,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) + 
#   facet_grid(~species)
# 
# ggplot(d, aes(biomassTo,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) + 
#   facet_grid(~species)

```

##Plot by spp,season
```{r, warning=FALSE}

# ggplot(d, aes(biomassFrom,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) + 
#   facet_grid(season~species)
# 
# ggplot(d, aes(biomassTo,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) + 
#   facet_grid(season~species)

```

##Plot by year, season
```{r, warning=FALSE}
#ggplot(filter(d), aes(biomassFrom,moved01PM, color=species)) +
#  geom_jitter(size=0.1,height = 0.15) +
#  stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) +
#  facet_grid(season~year)

# ggplot(filter(d,species=="bkt"), aes(biomassFrom,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) +
#   facet_grid(season~year, scales='free')
# 
# ggplot(filter(d,species=="bkt"), aes(biomassTo,moved01PM)) +
#   geom_jitter(size=0.1,height = 0.15) +
#   stat_smooth(method = "glm", method.args = list(family = "binomial"), se=F) +
#   facet_grid(season~year, scales='free')
```
Some interesting patterns in the above. Fish seem to move to high biomass sections. Maybe these are just good sections that can carry high biomass. Try scaling sections by average biomass over the study as a crude measure of habitat quality

```{r}

ggplot(filter(cdWB,area %in% c("inside","trib")), aes(as.factor(section), observedWeight)) +
  geom_boxplot(outlier.size = 0.2) + 
  scale_y_log10() +
  facet_grid(river+isYOY~season, scales = "free")

# means by isYOY,river,year,season,section
# this will be our 'raw' mean biomass and counts
byYRYSS <- cdWB %>% 
  filter(area %in% c("inside","trib")) %>%
  group_by(isYOY,river,year,season,section) %>%
  summarize( meanBiomassYRYSS = mean(observedWeight,na.rm = T),
             nYRYSS = n() ) %>%
  select(meanBiomassYRYSS,nYRYSS,isYOY,river,season,year,section)

ggplot(byYRYSS, aes(section, meanBiomassYRYSS, color = factor(year))) +
  geom_point() +
  geom_line() +  
  facet_grid(river+isYOY~season, scales = "free")

ggplot(byYRYSS, aes(section, nYRYSS, color = factor(year))) +
  geom_point() +
  geom_line() +  
  facet_grid(river+isYOY~season, scales = "free")



### run separate linear models for each isYOY,river,season
sectionModelB <- function(df) {
  lm( meanBiomassYRYSS ~ factor(section), data = df )
}

sectionModelN <- function(df) {
  lm( nYRYSS ~ factor(section), data = df )
}

# set up nesting
byYRYSS2 <- byYRYSS %>% 
  group_by(isYOY,river,season) %>%
  nest()

# run each model and add residuals
byYRYSS2 <- byYRYSS2 %>% 
  mutate( modelB = map(data, sectionModelB),
          residBiomass = map2(data, modelB, add_residuals),
          modelN = map(data, sectionModelN),
          residN = map2(data, modelN, add_residuals))

# to long format
rB <- unnest(byYRYSS2,residBiomass) %>% rename(residB = resid)
rN <- unnest(byYRYSS2,residN) %>% rename(residN = resid)

ggplot(rB, aes(section,residB, color = factor(year))) +
 geom_line() +
 facet_grid(river+isYOY~season, scales = "free")

ggplot(rN, aes(section,residN, color = factor(year))) +
 geom_line() +
 facet_grid(river+isYOY~season, scales = "free")





# 
# # means by isYOY,river,season,section - take out year
# # this is the variation, what we will merge YRS means back into
# byYRSS <- byYRYSS %>%
#   group_by(isYOY,river,season,section) %>%
#   summarize( meanBiomassYRSS = mean(meanBiomassYRYSS,na.rm = T),
#              meanN_YRSS = mean(nYRYSS,na.rm = T),
#              nYRSS = n()
#            )
# 
# ggplot(byYRSS, aes(section, meanBiomassYRSS)) +
#   geom_point(aes(color=isYOY)) + 
#   facet_grid(season ~ river)
# 
# # means by isYOY,river,season - take out section and year
# # this is the baseline
# byYRS <- byYRYSS %>%
#   group_by(isYOY,river,season) %>%
#   summarize( meanBiomassYRS = mean(meanBiomassYRYSS,na.rm = T),
#              sdBiomassYRS = sd(meanBiomassYRYSS,na.rm = T),
#              meanN_YRS = mean(nYRYSS,na.rm = T),
#              nYRS = n()
#            )
# 
# # merge means by YRS back in and get residuals
# r <- left_join(byYRSS,byYRS) %>%
#    mutate( residBiomass = meanBiomassYRSS - meanBiomassYRS,
#            residN = meanN_YRSS - meanN_YRS)
# 
# ggplot(r, aes(section, residBiomass)) +
#   geom_point(aes(color = isYOY)) + 
#   facet_grid(season ~ river, scales = 'free')
#  
# ggplot(r, aes(section, residN)) +
#   geom_point(aes(color = isYOY)) + 
#   facet_grid(season ~ river, scales = 'free')
 
```

##Merge biomass and N residuals back into cdWB
```{r}
#cdWB <- left_join(cdWB,select(r, isYOY,river,season,section,residBiomass,residN))
# merge in biomass and counts for the current sample 
cdWB <- left_join(cdWB,rB, by = c("river", "section", "year", "season", "isYOY"))
cdWB <- left_join(cdWB,select(rN, -meanBiomassYRYSS,-nYRYSS), by = c("river", "section", "year", "season", "isYOY"))

# merge in biomass and counts for the next sample
# need to rename residX so join doesn't rename variables by adding a .x or .y
rBNext <- rB %>% rename(residBNext = residB, meanBiomassYRYSSNext  = meanBiomassYRYSS, nYRYSSNext  = nYRYSS )
cdWB <- left_join(cdWB,rBNext, by = c("lagSection" = "section","river", "year", "season","isYOY")) 
        
rNNext <- rN %>% rename( residNNext = residN )
cdWB <- left_join(cdWB,select(rNNext, -meanBiomassYRYSS,-nYRYSS), by = c("lagSection" = "section","river", "year", "season","isYOY"))
        

ggplot(cdWB, aes(residB,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F)

summary( glm( moved01 ~ residB, family = "binomial", data = cdWB ) )

ggplot(cdWB, aes(residBNext,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F)

summary( glm( moved01 ~ residBNext, family = "binomial", data = cdWB ) )


ggplot(cdWB, aes(residN,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) 

summary( glm( moved01 ~ residN, family = "binomial", data = cdWB ) )

ggplot(cdWB, aes(residNNext,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) 

summary( glm( moved01 ~ residNNext, family = "binomial", data = cdWB ) )


```

