---
title: "Untitled"
output: html_document
---
```{r}
library(tidyverse)
library(lubridate)
library(getWBData)
```

```{r}
rm( list=setdiff(ls(), c("conDplyr","con")) )
#reconnect()
```


```{r read in data}

wb <- read.csv('wbCSV.csv', header = T, stringsAsFactors = F); wb$section <- as.numeric(wb$section)
ob <- read.csv('obCSV.csv', header = T, stringsAsFactors = F); #ob$section <- as.numeric(as.character(ob$section))
m <- read.csv('mCSV.csv', header = T, stringsAsFactors = F); #m$section <- as.numeric(as.character(m$section))
j <- read.csv('jCSV.csv', header = T, stringsAsFactors = F); j$quarter <- as.numeric(j$quarter) # there is a quarter "5 pool"

transects <- bind_rows(wb,ob,m,j)
transects$date <- as.Date(transects$date, "%m/%d/%Y")
transects$year <- year(transects$date)

# fix based on volumeDis data graphs
transects <- transects %>% filter( !(river == 'wb jimmy' & section == 5 & year == 2007 ))
transects$section <- ifelse(transects$river == 'wb jimmy' & transects$section < 5 & transects$year == 2007,
                            transects$section + 1, transects$section)

transects <- filter(transects,!(river == 'west brook' & date == '2009-10-19')) # this section was done at the end of a trib day and then repeated the next day. delete

transects$depth[transects$river == 'wb mitchell' & transects$depth > 50] <- NA
```

```{r calc volumes}

#transects$wD <- transects$width * transects$depth

vol1 <- transects %>%
  group_by(river,year,date,section,quarter) %>%
  summarize( sumDepths = sum(depth), n = n(), maxDepth = max(depth), width = max(width) )

vol <- vol1 %>%
  group_by(river,year,date,section) %>%
  mutate( volumeQ = sumDepths * width ) %>%
  summarize( volume = sum(volumeQ), nVol = n(), maxDepth = max(maxDepth), width = max(width) ) 
  

```

```{r raw data}

ggplot( vol, aes(section, volume, color = factor(year))) +
  geom_point() +
  geom_line() +
  facet_grid(~river)

ggplot( vol, aes(section, volume, color = factor(year))) +
  geom_point() +
  geom_line() +
  facet_grid(river~date)

```

Link in discharge for transect dates and scale either by 1) discharge or 2) z-score of discharge within a river/year
```{r scale to flow}
  reconnect() 

  flow <- data.frame(tbl(conDplyr, "data_daily_discharge")) %>%
    #filter(date %in% samples$date) %>%
    select(date, river, discharge)

  #scale by discharge
  vol <- left_join(vol,flow) %>% mutate( volumeDis = volume/(discharge + 1) )
  
  ggplot( vol, aes(section, volumeDis, color = factor(year))) +
    geom_point() +
    geom_line() +
    facet_wrap(~river, scales = 'free')
  
  #trends over time  
  ggplot( vol, aes(year, volumeDis, color = factor(section))) +
    geom_point() +
    geom_line() +
    facet_wrap(~river, scales = 'free')
  
  #####################################################################
  # scale by z-score
  meansByYear <- vol %>%
    group_by(river,year) %>%
    summarize( meanV = mean(volume), stdV = sd(volume) ) %>%
    ungroup()
  
  vol <- vol %>%
    left_join(.,meansByYear) %>%
    mutate( volumeZ = (volume - meanV)/stdV + 2 ) # + 2 so all values are > 0
  
  ggplot( vol, aes(section, volumeZ, color = factor(year))) +
    geom_point() +
    geom_line() +
    facet_wrap(~river, scales = 'free')

  #trends over time  
  ggplot( filter(vol, !is.na(date)), aes(year, volumeZ, color = factor(section))) +
    geom_point() +
    geom_line() +
    facet_wrap(~river, scales = 'free')
  

  
```

```{r, warning=FALSE, width = 9, height = 6}

  require(GGally)

  ggPairsData <- filter(ungroup(vol), !is.na(date)) %>% 
    select(river,section,year,volumeZ) %>%
    spread(year,volumeZ) %>%
    setNames( c('river', 'section', "a2007", "a2008", "a2009", "a2012", "a2015") )
  

  my_fn <- function(data, mapping, ...){
    p <- ggplot(data = data, mapping = mapping) + 
      geom_point() + 
      #geom_smooth(method=loess, fill="red", color="red", ...) +
      geom_smooth(method = lm, se = F, ...) +
      geom_text( aes(label = section), hjust = 0, vjust = 0 )
    p
  }
  
  riverTmp <- "wb jimmy"
  ggpairs( data = filter(ggPairsData, river == riverTmp),
           col = 3:7,
           lower = list(continuous = my_fn),
           title = riverTmp
  )
  
  ggpairs( data = filter(ggPairsData),
           col = 3:7,
           ggplot2::aes(color = river, alpha = 0.6),
           lower = list(continuous = my_fn)
  )

  
```

Steps:

Identify sections with big volume deltas based on transects
Can we use changes in biomass between transect samples to pinpoint when a section volume changed?



Get means by section
```{r}

  meansBySection <- vol %>%
    group_by(river,section) %>%
    summarize( meanVolumeDis = mean(volumeDis),
               meanVolumeZ = mean(volumeZ),
               meanVolumeBySection = mean(volume),
               meanMaxDepthBySection = mean(maxDepth),
               minYear = min(year))

```


Linearly interpolate missing years
```{r}
library(zoo)


vol <- vol %>% 
  group_by(river,section) %>%
  complete(nesting(river,section), year = 1997:2015 ) %>%
  mutate( volumeDis = na.approx(volumeDis, rule = 2),
          volumeZ =   na.approx(volumeZ, rule = 2),
          volumeNAApprox = na.approx(volume, rule = 2),
          maxDepthNAApprox = na.approx(maxDepth, rule = 2))  

vol <- vol %>%
  left_join(.,meansBySection) 

vol$volumeDis <- ifelse( vol$year < vol$minYear, vol$meanVolumeDis, vol$volumeDis )
vol$volumeZ   <- ifelse( vol$year < vol$minYear, vol$meanVolumeZ, vol$volumeZ )
vol$volumeNAApprox   <- ifelse( vol$year < vol$minYear, vol$meanVolumeBySection, vol$volumeNAApprox )
vol$maxDepthNAApprox <- ifelse( vol$year < vol$minYear, vol$meanMaxDepthBySection, vol$maxDepthNAApprox )


 # mutate( volumeDis = ifelse( year < 2008, mean(volumeDis), na.approx(volumeDis, rule = 2)),
#          volumeZ = ifelse( year < 2008, mean(volumeZ), na.approx(volumeZ, rule = 2)) ) 

  ggplot( vol, aes(section, volumeZ, color = factor(year))) +
    geom_point( aes(shape = is.na(date)) ) +
    geom_line() +
    facet_wrap(~river, scales = 'free')
 
  ggplot( vol, aes(section, volumeNAApprox, color = factor(year))) +
    geom_point( aes(shape = is.na(date)) ) +
    geom_line() +
    facet_wrap(~river, scales = 'free') 
   
    #trends over time  
  ggplot( filter(vol,river == "west brook"), aes(year, volumeZ, color = factor(section))) +
    geom_point( aes(shape = is.na(date)) ) +
    geom_line() +
    facet_wrap(~section, scales = 'free') +
    theme(strip.background = element_blank(),
          strip.text.x = element_blank())

```


```{r}

volMeansByRiver <- vol %>%
  group_by(river) %>%
  summarize( meanMaxDepthByRiver = mean(maxDepth, na.rm = T),
             meanVolByRiver = mean(volume, na.rm = T) )

vol <- left_join( vol, volMeansByRiver ) %>% mutate( pool01 = 1*(volumeNAApprox > meanVolByRiver & maxDepthNAApprox > meanMaxDepthByRiver )  )

ggplot( vol, aes(volume,maxDepth,color = factor(pool01)) ) +
  geom_point() +
  geom_vline( aes(xintercept = meanVolByRiver), data = volMeansByRiver ) +
  geom_hline( aes(yintercept = meanMaxDepthByRiver), data = volMeansByRiver ) +
  facet_wrap(~river, scales = "free")

ggplot( vol, aes(volumeNAApprox,maxDepthNAApprox,color = factor(pool01)) ) +
  geom_point() +
  geom_vline( aes(xintercept = meanVolByRiver), data = volMeansByRiver ) +
  geom_hline( aes(yintercept = meanMaxDepthByRiver), data = volMeansByRiver ) +
  facet_wrap(~river, scales = "free") +
  theme(legend.position = "none")

#trends over time  
ggplot( filter(vol,river == "west brook"), aes(year, pool01, color = factor(section))) +
    geom_point( aes(shape = is.na(date)) ) +
    geom_line() +
    facet_wrap(~section) +
    theme(legend.position = "none")

```



```{r}
# means by isYOY,river,year,season,section
# this will be our 'raw' mean biomass and counts

load('cdWBForTransects.RData')

cdWB <- left_join(cdWB,vol, by = c("river", "section", "year"))

byYRYSS <- cdWB %>% 
  group_by(isYOY,river,year,season,section,species) %>%
  summarize( meanBiomassYRYSS = mean(observedWeight,na.rm = T),
             meanBiomassByVolZYRYSS = mean(observedWeight/(volumeZ), na.rm = T), 
             varBiomassByVolZYRYSS = var(observedWeight/(volumeZ), na.rm = T), 
             nYRYSS = n(),
             meanVolZ = mean(volumeZ, na.rm = T),
             sumBiomassYRYSS = sum(observedWeight,na.rm = T)) %>%
  select(meanBiomassYRYSS,nYRYSS,meanBiomassByVolZYRYSS,varBiomassByVolZYRYSS,meanVolZ,sumBiomassYRYSS,isYOY,river,season,year,section,species)

cdWB <- left_join(cdWB,byYRYSS)

```

```{r}
riverTmp <- "west brook"
isYOYTmp <- 0
seasonTmp <-  3

ggplot(filter(byYRYSS, river == riverTmp & isYOY == isYOYTmp & season == seasonTmp), aes(section, meanBiomassYRYSS, color = factor(year))) +
  geom_point( aes(size = nYRYSS) ) +
  geom_line() + 
 # ylim(-100,100) +
  facet_wrap(~year)#, scales = "free")

ggplot(filter(byYRYSS, river == riverTmp & isYOY == isYOYTmp & season == seasonTmp), aes(section, meanVolZ, color = factor(year))) +
  geom_point( ) +
  geom_line() + 
 # ylim(-100,100) +
  facet_wrap(~year)#, scales = "free")

ggplot(filter(byYRYSS, river == riverTmp & isYOY == isYOYTmp & season == seasonTmp), aes(section, meanBiomassByVolZYRYSS, color = factor(year))) +
  geom_point( aes(size = nYRYSS) ) +
  geom_line() + 
 # ylim(-100,100) +
  facet_wrap(~year)#, scales = "free")

sectionTmp <- 6
ggplot(filter(byYRYSS, river == riverTmp & isYOY == isYOYTmp & section == sectionTmp), aes(year*(season/4), meanBiomassYRYSS, color = factor(year))) +
  geom_point( aes(size = nYRYSS) ) +
  geom_line() + 
 # ylim(-100,100) +
  facet_wrap(~year)#, scales = "free")
```

SumOfBiomass consitent over years?
```{r}
riverTmp2 <- "wb jimmy"
isYOYTmp2 <- 0
seasonTmp2 <-  3
speciesTmp2 <- "bkt"

ggplot(filter(byYRYSS, river == riverTmp2 & isYOY == isYOYTmp2 & season == seasonTmp2 &  species == speciesTmp2), aes(year , sumBiomassYRYSS, color = factor(section))) +
  geom_point( aes(size = nYRYSS)) +
  geom_line() + 
  ggtitle(paste0(riverTmp2,", ","isYOY = ",isYOYTmp2,", ",'season = ',seasonTmp2,", ",speciesTmp2)) +
  facet_wrap(~section)#, scales = "free")


```


Look for relationships btw vol and biomass
```{r}
ggplot( filter( byYRYSS, year %in% c(2007,2008,2009,2012,2015) & season == 3 & isYOY == 0 & species == 'bkt'), aes(meanVolZ,meanBiomassYRYSS)) +
  geom_point( aes( size = nYRYSS) ) +
  geom_smooth( method = 'lm', se = F) +
  facet_grid(season + year ~ river, scales = 'free')

ggplot( filter( byYRYSS, year %in% c(2007,2008,2009,2012,2015) & season == 3 & isYOY == 0 & year == 2015), aes(meanVolZ,meanBiomassYRYSS)) +
  geom_point( aes( size = nYRYSS) ) +
  geom_smooth( method = 'lm', se = F) +
  facet_grid(season + species ~ river, scales = 'free')

ggplot( filter( byYRYSS, year %in% c(2007,2008,2009,2012,2015)), aes(meanVolZ,varBiomassByVolZYRYSS )) +
  geom_point() +
  facet_grid(season~river, scales = 'free')
```


```{r}

ggplot(filter(cdWB), aes(volumeZ,moved01)) + # aes(meanBiomassByAreaYRYSS,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-100,100)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(season~species, scales = "free")

ggplot(filter(cdWB), aes(volumeZ,moved01PM)) + # aes(meanBiomassByAreaYRYSS,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-100,100)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(season~species, scales = "free")

ggplot(filter(cdWB), aes(meanBiomassByVolZYRYSS,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-100,100)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(season~species, scales = "free")

ggplot(filter(cdWB), aes(meanBiomassByVolZYRYSS,moved01PM)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-100,100)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(season~species, scales = "free")

ggplot(filter(cdWB, moveDir >= 0), aes(meanBiomassByVolZYRYSS,moveDir)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-100,100)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(season~species, scales = "free")
ggplot(filter(cdWB, moveDir <= 0)%>% mutate(moveDir = moveDir* -1), aes(meanBiomassByVolZYRYSS,moveDir)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-100,100)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(season~species, scales = "free")

```

Divide up by transitions among rivers ('trans')
```{r}

ggplot(filter(cdWB,species == "bkt" & as.numeric(riverOrdered) == lagRiver & meanBiomassByVolZYRYSS < 75), aes(meanBiomassByVolZYRYSS,moved01)) +
  geom_jitter(size = 0.1,height = 0.15) +
  #xlim(c(-100,100)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  facet_grid(season ~ trans, scales = "free")


```

Growth rate consequences of moving
```{r}

```



```{r save data}

save(transects, vol, meansBySection, meansByYear, file = 'transects.RData')

```

